Computation and comparison of the stable Northeastern US marine boundary layer

### 2020-12-03 Paper/main.tex
Finished making Alan's changes

\section{Nomenclature}
\fxnote{Update symbols}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
LES       & Large Eddy Simulation \\
$u_\tau$   & Friction velocity    \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy
\end{longtable*}}

@@

\section{Conclusion}
In this study we have simulated the stable offshore boundary layer
matching the characteristics of measurements from Nantucket Sound in
the Northeastern US.  Three cases with wind speeds of 5 m/s, 10 m/s,
and 15 m/s were selected, and simulations were performed using the
ExaWind codes to match the TI and stability levels observed at from
the Cape Wind platform.  When examining the effect of stratification
using the Obukhov length, differences in the ABL behavior were seen
bewteen the very stable boundary layer cases (5 m/s and 10 m/s), and
the stably stratified case (15 m/s).  The very stable cases had
shorter turbulent structures and less growth in the boundary layer
compared to the 15 m/s case.  The behavior of the three stable
boundary layers was also compared to the neutral and unstable cases
studied in \cite{cheung2020large}.  As expected, the turbulent
structures in all stable cases were noticeably smaller than in the
neutral and unstable boundary layers.

>>>>

\section{Nomenclature}
\fxnote{Update symbols}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
LES       & Large Eddy Simulation \\
$u_\tau$   & Friction velocity    \\
$\rho$    & Density \\
$\rho_\circ$ & Reference density  \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$\theta$  & Potential temperature \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy
\end{longtable*}}

@@

\section{Conclusion}
In this study we have simulated the stable offshore boundary layer
matching the characteristics of measurements from Nantucket Sound in
the Northeastern US.  Three cases with wind speeds of 5 m/s, 10 m/s,
and 15 m/s were selected, and simulations were performed using the
ExaWind codes to match the TI and stability levels observed at from
the Cape Wind platform.  When examining the effect of stratification
using the Obukhov length, differences in the ABL behavior were seen
between the very stable boundary layer cases (5 m/s and 10 m/s), and
the stably stratified case (15 m/s).  The very stable cases had
shorter turbulent structures and less growth in the boundary layer
compared to the 15 m/s case.  The behavior of the three stable
boundary layers was also compared to the neutral and unstable cases
studied in \cite{cheung2020large}.  As expected, the turbulent
structures in all stable cases were noticeably smaller than in the
neutral and unstable boundary layers.

### 2020-12-03 Paper/methodology.tex
Fix typos and stuff
Finished making Alan's changes

From the measured distribution of atmospheric stabilities, turbulent
kinetic energies (TKE), and wind speeds from the Cape Wind platform at
z=20m, a stable set of conditions were chosen as targets for this
computational study.  A summary of the conditions for all stability
classes is given table \ref{tab:CapeWindMeasurements}. In this work we
focus on the three stable atmospheric conditions at wind speeds of
5m/s, 10m/s, and 15m/s, while the neutral and unstable conditions
where studied in previous work \cite{cheung2020large}.

@@

Large Eddy Simulations (LES) of the stable atmopsheric conditions described in
Table~\ref{tab:CapeWindMeasurements} are performed using the open-source,
ExaWind simulation environment\footnote{\url{https://github.com/exawind}}.
ExaWind is a suite of high-fidelity modeling tools for analyzing complex
microscale atmospheric boundary layer flows and wind farm flows. The present
study uses two different incompressible computational fluid dynamic solvers
available within the ExaWind suite --
Nalu-Wind~\footnote{\url{https://github.com/exawind/nalu-wind}} and
AMR-Wind~\footnote{\url{https://github.com/exawind/amr-wind}}. This section
briefly describes the incompressible Navier-Stokes equations solved in the
solvers, the specifics of the turbulence model and the wall shear stress
formulation used to model the terrain, and the details of the numerical
discretization of the two equations in the two codes.

@@

\subsubsection{Governing equations}
Both CFD codes solve the incompressible form of Navier-Stokes equations with
appropriate models for turbulence closure. Equation~\ref{eqn:ns-les} shows the
continuity and momentum equations with all the terms necessary to model the
atmospheric boundary layer problem. Term $\mathbf{I}$ represents the pressure
gradient forces as a deviation from hydrostatic and horizontal mean gradient,
term $\mathbf{II}$ represents the contribution from viscous and sub-filter scale
stresses, term $\mathbf{III}$ represents the contribution from Coriolis forces
due to earth's rotation, term $\mathbf{IV}$ represents the effects of buoyancy,
and term $\mathbf{V}$ represents source terms necessary to drive the flow to a
desired horizontal mean velocity at prescribed heights.

@@

\begin{align*}
  \frac{\rho}{\rho_\circ} \approx 1 - \beta \left( \theta - \theta_\circ \right)
\end{align*}
This leads to a forcing term that depends on potential temperature ($\theta$),
the reference density $\rho_\circ = \bar{\rho}$, and thermal expansion
coefficient $\beta \approx 1 / \theta_\circ$ as
\begin{align}
  \left(\rho - \rho_\circ \right) g_i = -\rho_\circ\, \beta\, g_i \left( \theta - \theta_\circ \right)
\end{align}

For ABL problems, the conservation of energy equation is usually written in
terms of potential temperature as shown in Eq.~\ref{eqn:pot-temp-les}.
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{\theta}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{\theta} \right) = - \frac{\partial}{\partial x_j} \hat{q}_j \label{eqn:pot-temp-les}
\end{align}
Under the assumption of ideal gas conditions and constant specific heat capacity
$c_p$, the gradients in potential temperature are proportional to the gradients
in absolute temperature, i.e.,
\begin{align*}
   \left[ \frac{\partial T}{\partial t}, \frac{\partial T}{\partial x}, \frac{\partial T}{\partial y} \right] =
   \left( \frac{\bar{p}}{p_\circ} \right)^{\left(\frac{R}{c_p}\right)} \left[ \frac{\partial \theta}{\partial t}, \frac{\partial \theta}{\partial x}, \frac{\partial \theta}{\partial y} \right]
\end{align*}
Furthermore, ignoring the pressure and viscous work terms in
Eq.~\ref{eqn:pot-temp-les}, for incompressible flows, it can be shown that
solving the enthalpy equation is equivalent to solving the potential temperature
equation. The enthalpy equation solved in Nalu-Wind for wind energy problems is
shown below
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{h}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{h} \right) = - \frac{\partial}{\partial x_j} q_j \label{eqn:enthalpy-les}
\end{align}
It is noted here that the terms $\hat{q}_j$ (Eq.~\ref{eqn:pot-temp-les}) and
$q_j$ (Eq.~\ref{eqn:enthalpy-les}) are not equivalent and must be scaled
appropriately. User can still provide the appropriate initial and boundary
conditions in terms of potential temperature field. Under these assumptions and
conditions, the resulting solution can then be interpreted as the variation of
potential temperature field in the computational domain.

The subgrid-scale kinetic energy ($k$) one-equation turbulence model
(Eq.~\ref{eqn:ksgs-les}), based on Moeng~\cite{Moeng1984}, is used for
LES turbulence closure.

\begin{align}
  \frac{\partial}{\partial t}\left(\bar{\rho} k \right) + \frac{\partial}{\partial x_j} \bar{\rho} k \widetilde{u_j} &= \frac{\partial}{\partial x_j} \left (2 \nu_t \frac{\partial k}{\partial x_j} \right) - 2 \nu_t \widetilde{S_{ij}} \widetilde{S_{ij}} + \frac{g}{\theta_\circ} \tau_{\theta_i} - C_{\epsilon} \frac{k^{3/2}}{l} \label{eqn:ksgs-les}
\end{align}
where
\begin{align*}
  C_\epsilon &= 0.93 & \nu_t &= C_\epsilon l \sqrt{k} \\
  \tau_{\theta i} &= - 2\ \frac{\nu_t}{\mathrm{Pr}_t}\ \frac{\partial \widetilde{\theta}}{\partial x_i} & \mathrm{Pr}_t &= \left( 1 + \frac{2l}{\Delta} \right)^{-1} \\
\end{align*}
The choice of the length scale ($l$) depends on whether the stratification is positive
or negative, based on the recommendation by Moeng~\cite{Moeng1984}, and is given
by
\begin{align*}
  l =
  \begin{cases}
    \Delta = \sqrt[3]{\Delta_x \Delta_y \Delta_z}, & \frac{\partial \widetilde{\theta}}{\partial z} < 0 \\
    0.76\, \sqrt{k}\, \frac{g}{\theta_\circ}\, \left( \frac{\partial \widetilde{\theta}}{\partial z}\right)^{-1/2}, &  \frac{\partial \widetilde{\theta}}{\partial z} > 0 \\
  \end{cases}
\end{align*}

\begin{comment}
  \fxnote{Shreyas write this part.  Currently placeholder equations below.}\\

  The low-Mach number equations:
  \begin{align}
    \frac{\partial \rho} {\partial t} + \frac{\partial \rho u_j}{\partial x_j} & = 0, \\
    \frac{\partial \rho u_i}{\partial t} + \frac{\partial \rho u_j u_i}{\partial x_j}
    + \frac{\partial P}{\partial x_i} & = \frac{\partial \tau_{ij}}{\partial x_j}
                                        + \left( \rho - \rho_{\circ} \right) g_i, \\
    \frac{\partial \rho h}{\partial t} + \frac{\partial \rho u_j h}{\partial x_j} & =
                                                                                    - \frac{\partial q_j}{\partial x_j} + \frac{\partial P_{th}}{\partial t},
  \end{align}
\end{comment}

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{S_1 \average{\widetilde{u}_1} + \average{S_1} (\widetilde{u}_1 - \average{\widetilde{u}_1} ) }{ \average{S_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{S_1 \average{\widetilde{v}_1} + \average{S_1} (\widetilde{v}_1 - \average{\widetilde{v}_1} ) }{ \average{S_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{S_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{S_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{S_1} ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $S_1$ is the resolved wind speed at the first grid level and
$\average{f}$ represents the horizontal planar average of $f$. The mean
sub-filter stresses are computed using Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{\widetilde{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{S_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{S_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{S_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.

@@

Nalu-Wind~\cite{SpragueAVR2020} is an open-source, wind-focused fork of the Nalu CFD
code~\cite{Domino:2015}. The codebase heavily leverages the open-source Trilinos
suite of numerical libraries~\cite{Heroux:2003}. Nalu-Wind uses an
unstructured-grid, node-centered, finite-volume formulation for solving the incompressible
Navier-Stokes equations. Two spatial discretization approaches are supported
within the codebase -- the edge-based scheme and control-volume finite-element
scheme. The edge-based scheme, used in the present study, is similar to the
classical cell-centered finite volume formulation formulated on a dual volume
surrounding the nodes of the mesh. The governing equations are discretized in
time using a split-operator approach. Implicit backward difference formula
(BDF2) timestepping algorithm with an approximate pressure projection
scheme~\cite{Moen-Domino:2003} is used to advance the equations in time.
Multiple Picard-style outer-loop iterations are employed within each timestep
to minimize the errors introduced from split-operator approximations. Nalu-Wind
comes equipped with a suite of Reynolds-Averaged Navier-Stokes (RANS), LES, and
Detached Eddy Simulation (DES) turbulence models to resolve the disparate
spatial and temporal scales encountered when modeling wind plant flows.

Nalu-Wind heavily utilizes the Sierra Toolkit (STK) library~\cite{Edwards:2010}
for managing the unstructured mesh data structures. The linear systems resulting
from the discretized equations are solved using solvers available within the
Belos and MueLu libraries within Trilinos. The momentum and scalar transport
equations are solved using GMRES iterative solver with symmetric Gauss-Seidel
(SGS) preconditioners. The elliptic pressure-Poisson equation is solved using
GMRES solvers with a smoothed-aggregated algebraic multigrid preconditioner. The
reader is referred to Sprague et al.~\cite{SpragueAVR2020} and
Domino~\cite{Domino:2015} more more details.

@@

As part of a suite of physics codes within the open-source project called ExaWind,
AMR-Wind enables more efficient and scalable simulations of wind power plants.
This is because it is built on top of the AMReX software framework which contains
all of the functionality needed to develop massively parallel, block-structured
adaptive mesh refinement (AMR) applications \cite{AMReX_JOSS}.
Since AMR-Wind is limited to Cartesian block-structured grids it is specialized
to handle atmospheric boundary layer physics and wind turbine wakes.
The block-structured grids enable more efficient algorithms such as
Multi-Level Multi-Grid (MLMG) \cite{AMReX_JOSS}
and are well suited for next generation supercomputers that use GPU's.

AMR-Wind solves the incompressible Navier-Stokes equations and is generalized to
handle variable density and viscosity. Additionally scalar transport equations can be
solved such as potential temperature or turbulence models. The discretization in AMR-Wind
is based on the approximate projection method used in IAMR \cite{almgren1998conservative}
and incflo \cite{sverdrup2018highly}. It is a semi-staggered scheme where the velocity and scalar
variables are located at cell centers and pressure is located at nodes. Pressure is also staggered
in time so that pressure and the pressure gradient (located at cell centers) are at time $n+1/2$.
The time discretization is handled with a Crank Nicholson approach where the right hand side
of the system of equations is evaluated at time $n+1/2$. The advection term is handled explicitly
using an upwind finite-volume method. There are multiple options
within AMR-Wind, but in this paper only the high-order Godunov Piecewise Parabolic
method (PPM) is used \cite{Colella1984}. The advection terms are projected to face
centers at $n+1/2$ and are corrected using an exact MAC projection \cite{almgren1998conservative}
which guarantees a divergence free flow. The diffusion terms can be handled explicitly, semi-implicitly,
or fully implicit and are the discretized using a second-order central difference formula.
For the simulations in this paper we use a fully implicit scheme as the variable viscosity
from the eddy viscosity may cause time step restrictions. After the scalar equations
and the momentum equations are advanced in time a nodal projection is used to
approximately correct the velocity field to make it divergence free. Where the nodal projection
is based on a node centered finite element method \cite{almgren1998conservative}.

@@

The computational methodology for the AMR-Wind and Nalu-Wind LES codes
follows practices similar to the previous offshore ABL study, with
some modifications to handle the stable stratification.  As in
\cite{cheung2020large}, the domain was a square prism geometry, with
the $x$ and $y$ coordinates aligned in the East and North directions,
respectively.  The domain size and mesh requirements were determined
through a numerical grid study.  As discussed in section
\ref{sec:gridstudy}, the horizontal dimensions of 750m $\times$ 750m
were found to be sufficient to capture any large scale structures in
the boundary layer, and in all cases the vertical dimension of 1000m
was also used.  The grid study in section \ref{sec:gridstudy} also
showed that the stable stratification required greater mesh refinement
to capture the smaller turbulent scales, and a uniform cell size of
2.5m was adopted for the offshore LES computations across both codes.

@@

\subsubsection{Determination of surface roughness and prescribed temperature }
To match the measured TI conditions given in table
\ref{tab:CapeWindMeasurements}, the surface roughness and prescribed
surface temperature change were adjusted through an initial
trial-and-error process.  The final surface roughness of $z_0$=0.0005m
used the stable offshore conditions (see table \ref{tab:z0tempparam})
matched the roughness used in the neutral 15m/s and unstable 5m/s and
10m/s cases from the previous study \cite{cheung2020large}.  This
value of surface roughness is consistent with the measurements from
the North Sea \cite{taylor2001dependence}, which found values of $z_0$
ranging from $5 \times 10^{-5}$m to $5\times 10^{-3}$m.

The values of the prescribed temperature change at the ocean surface
is also given in table \ref{tab:CapeWindMeasurements}.  As the wind
speed increases, a larger temperature decrease was used to maintain
similar levels of stable stratification, as expected from
Monin-Obukhov similarity theory.

>>>>

From the measured distribution of atmospheric stabilities, turbulent
kinetic energies (TKE), and wind speeds from the Cape Wind platform at
z=20m, a stable set of conditions were chosen as targets for this
computational study.  A summary of the conditions for all stability
classes is given in table \ref{tab:CapeWindMeasurements}. In this work we
focus on the three stable atmospheric conditions at wind speeds of
5m/s, 10m/s, and 15m/s, while the neutral and unstable conditions
where studied in previous work \cite{cheung2020large}.

@@

Large Eddy Simulations of the stable atmopsheric conditions described in
Table~\ref{tab:CapeWindMeasurements} are performed using the open-source,
ExaWind simulation environment\footnote{\url{https://github.com/exawind}}.
ExaWind is a suite of high-fidelity modeling tools for analyzing complex
microscale atmospheric boundary layer flows and wind farm flows. The present
study uses two different incompressible computational fluid dynamic solvers
available within the ExaWind suite,
Nalu-Wind~\footnote{\url{https://github.com/exawind/nalu-wind}} and
AMR-Wind~\footnote{\url{https://github.com/exawind/amr-wind}}. This section
briefly describes the incompressible Navier-Stokes equations solved,
the specifics of the turbulence model and the wall shear stress
formulation used to model the terrain, and the details of the numerical
discretization in the two codes.

@@

\subsubsection{Governing equations}
Both CFD codes solve the incompressible form of the Navier-Stokes equations with
appropriate models for turbulence closure. Equation~\ref{eqn:ns-les} shows the
continuity and momentum equations with all the terms necessary to model the
atmospheric boundary layer problem. Term $\mathbf{I}$ represents the pressure
gradient forces as a deviation from hydrostatic and horizontal mean gradients,
term $\mathbf{II}$ represents the contribution from viscous and sub-filter scale
stresses, term $\mathbf{III}$ represents the contribution from Coriolis forces
due to earth's rotation, term $\mathbf{IV}$ represents the effects of buoyancy,
and term $\mathbf{V}$ represents source terms necessary to drive the flow to a
desired horizontal mean velocity at prescribed heights.

@@

\begin{align}
  \frac{\rho}{\rho_\circ} \approx 1 - \beta \left( \theta - \theta_\circ \right)
\end{align}
This leads to a forcing term that depends on potential temperature ($\theta$),
the reference density $\rho_\circ = \bar{\rho}$, and thermal expansion
coefficient $\beta \approx 1 / \theta_\circ$ as
\begin{align}
  \left(\rho - \rho_\circ \right) g_i = -\rho_\circ\, \beta\, g_i \left( \theta - \theta_\circ \right)
\end{align}

For ABL problems, the conservation of energy equation is usually written in
terms of potential temperature as shown in Eq.~\ref{eqn:pot-temp-les}.
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{\theta}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{\theta} \right) = - \frac{\partial}{\partial x_j} \hat{q}_j \label{eqn:pot-temp-les}
\end{align}
Under the assumption of ideal gas conditions and constant specific heat capacity
$c_p$, the gradients in potential temperature are proportional to the gradients
in absolute temperature, i.e.,
\begin{align}
   \left[ \frac{\partial T}{\partial t}, \frac{\partial T}{\partial x}, \frac{\partial T}{\partial y} \right] =
   \left( \frac{\bar{p}}{p_\circ} \right)^{\left(\frac{R}{c_p}\right)} \left[ \frac{\partial \theta}{\partial t}, \frac{\partial \theta}{\partial x}, \frac{\partial \theta}{\partial y} \right]
\end{align}
Furthermore, ignoring the pressure and viscous work terms in
Eq.~\ref{eqn:pot-temp-les}, for incompressible flows, it can be shown that
solving the enthalpy equation is equivalent to solving the potential temperature
equation. The enthalpy equation solved in Nalu-Wind for wind energy problems is
shown below
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{h}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{h} \right) = - \frac{\partial}{\partial x_j} q_j . \label{eqn:enthalpy-les}
\end{align}
It is noted here that the terms $\hat{q}_j$ (Eq.~\ref{eqn:pot-temp-les}) and
$q_j$ (Eq.~\ref{eqn:enthalpy-les}) are not equivalent and must be scaled
appropriately. User can still provide the appropriate initial and boundary
conditions in terms of potential temperature field. Under these assumptions and
conditions, the resulting solution can then be interpreted as the variation of
potential temperature field in the computational domain.

The subgrid-scale kinetic energy ($k$) one-equation turbulence model
(Eq.~\ref{eqn:ksgs-les}), based on Moeng~\cite{Moeng1984}, is used for
LES turbulence closure,

\begin{align}
  \frac{\partial}{\partial t}\left(\bar{\rho} k \right) + \frac{\partial}{\partial x_j} \bar{\rho} k \widetilde{u_j} &= \frac{\partial}{\partial x_j} \left (2 \nu_t \frac{\partial k}{\partial x_j} \right) - 2 \nu_t \widetilde{S_{ij}} \widetilde{S_{ij}} + \frac{g}{\theta_\circ} \tau_{\theta_i} - C_{\epsilon} \frac{k^{3/2}}{l} \label{eqn:ksgs-les}
\end{align}
where
\begin{align}
  C_\epsilon &= 0.93 & \nu_t &= C_\epsilon l \sqrt{k} \\
  \tau_{\theta i} &= - 2\ \frac{\nu_t}{\mathrm{Pr}_t}\ \frac{\partial \widetilde{\theta}}{\partial x_i} & \mathrm{Pr}_t &= \left( 1 + \frac{2l}{\Delta} \right)^{-1} \\
\end{align}
The choice of the length scale ($l$) depends on whether the stratification is positive
or negative, based on the recommendation by Moeng~\cite{Moeng1984}, and is given
by
\begin{align}
  l =
  \begin{cases}
    \Delta = \sqrt[3]{\Delta_x \Delta_y \Delta_z}, & \frac{\partial \widetilde{\theta}}{\partial z} < 0 \\
    0.76\, \sqrt{k}\, \frac{g}{\theta_\circ}\, \left( \frac{\partial \widetilde{\theta}}{\partial z}\right)^{-1/2}, &  \frac{\partial \widetilde{\theta}}{\partial z} > 0 \\
  \end{cases}
\end{align}

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{S_1 \average{\widetilde{u}_1} + \average{S_1} (\widetilde{u}_1 - \average{\widetilde{u}_1} ) }{ \average{S_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{S_1 \average{\widetilde{v}_1} + \average{S_1} (\widetilde{v}_1 - \average{\widetilde{v}_1} ) }{ \average{S_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{S_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{S_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{S_1} ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $S_1$ is the resolved wind speed at the first grid level and
$\average{\bullet}$ represents the horizontal planar average. The mean
sub-filter stresses are computed using Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{\widetilde{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{S_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{S_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{S_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.

@@

Nalu-Wind~\cite{SpragueAVR2020} is an open-source, wind-focused fork of the Nalu CFD
code~\cite{Domino:2015}. The codebase heavily leverages the open-source Trilinos
suite of numerical libraries~\cite{Heroux:2003}. Nalu-Wind uses an
unstructured-grid, node-centered, finite-volume formulation for solving the incompressible
Navier-Stokes equations. Two spatial discretization approaches are supported
within the codebase: the edge-based scheme and control-volume finite-element
scheme. The edge-based scheme, used in the present study, is similar to the
classical cell-centered finite volume formulation formulated on a dual volume
surrounding the nodes of the mesh. The governing equations are discretized in
time using a split-operator approach. Implicit backward difference formula
(BDF2) timestepping algorithm with an approximate pressure projection
scheme~\cite{Moen-Domino:2003} is used to advance the equations in time.
Multiple Picard-style outer-loop iterations are employed within each timestep
to minimize the errors introduced from split-operator approximations. Nalu-Wind
comes equipped with a suite of Reynolds-Averaged Navier-Stokes (RANS), LES, and
Detached Eddy Simulation (DES) turbulence models to resolve the disparate
spatial and temporal scales encountered when modeling wind plant flows.

Nalu-Wind heavily utilizes the Sierra Toolkit (STK) library~\cite{Edwards:2010}
for managing the unstructured mesh data structures. The linear systems resulting
from the discretized equations are solved using solvers available within the
Belos and MueLu libraries within Trilinos. The momentum and scalar transport
equations are solved using GMRES iterative solvers with symmetric Gauss-Seidel
(SGS) preconditioners. The elliptic pressure-Poisson equation is solved using
GMRES solvers with a smoothed-aggregated algebraic multigrid preconditioner. The
reader is referred to Sprague et al.~\cite{SpragueAVR2020} and
Domino~\cite{Domino:2015} for more details.

@@

As part of a suite of physics codes within the open-source project called ExaWind,
AMR-Wind enables more efficient and scalable simulations of wind power plants.
This is because it is built on top of the AMReX software framework which contains
all of the functionality needed to develop massively parallel, block-structured
adaptive mesh refinement (AMR) applications \cite{AMReX_JOSS}.
Since AMR-Wind is limited to Cartesian block-structured grids, it is specialized
to handle ABL physics and wind turbine wakes.
The block-structured grids enable more efficient algorithms such as
Multi-Level Multi-Grid (MLMG) \cite{AMReX_JOSS}
and are well suited for next generation supercomputers that use GPU's.

AMR-Wind solves the incompressible Navier-Stokes equations and is generalized to
handle variable density and viscosity. Additionally scalar transport equations can be
solved such as potential temperature or turbulence models. The discretization in AMR-Wind
is based on the approximate projection method used in IAMR \cite{almgren1998conservative}
and incflo \cite{sverdrup2018highly}. It is a semi-staggered scheme where the velocity and scalar
variables are located at cell centers and pressure is located at nodes. Pressure is also staggered
in time so that pressure and the pressure gradient (located at cell centers) are at time $n+1/2$.
The time discretization is handled with a Crank Nicholson approach where the right hand side
of the system of equations is evaluated at time $n+1/2$. The advection term is handled explicitly
using an upwind finite-volume method. There are multiple options
within AMR-Wind, but in this paper only the high-order Godunov Piecewise Parabolic
method (PPM) is used \cite{Colella1984}. The advection terms are projected to face
centers at $n+1/2$ and are corrected using an exact MAC projection \cite{almgren1998conservative}
which guarantees a divergence free flow. The diffusion terms can be handled explicitly, semi-implicitly,
or fully implicit and are discretized using a second-order central difference formula.
For the simulations in this paper, we use a fully implicit scheme as the variable viscosity
from the eddy viscosity may cause time step restrictions. After the scalar equations
and the momentum equations are advanced in time, a nodal projection is used to
approximately correct the velocity field to make it divergence free.  The nodal projection
is based on a node centered finite element method \cite{almgren1998conservative}.

@@

The computational methodology for the AMR-Wind and Nalu-Wind LES codes
follows practices similar to the previous offshore ABL study \cite{cheung2020large}, with
some modifications to handle the stable stratification.  As in
\cite{cheung2020large}, the domain was a square prism geometry, with
the $x$ and $y$ coordinates aligned in the East and North directions,
respectively.  The domain size and mesh requirements were determined
through a numerical grid study.  As discussed in section
\ref{sec:gridstudy}, the horizontal dimensions of 750m $\times$ 750m
were found to be sufficient to capture any large scale structures in
the boundary layer, and in all cases the vertical dimension of 1000m
was also used.  The grid study in section \ref{sec:gridstudy} also
showed that the stable stratification required greater mesh refinement
to capture the smaller turbulent scales, and a uniform cell size of
2.5m was adopted for the offshore LES computations across both codes.

@@

\subsubsection{Determination of surface roughness and prescribed temperature }
To match the measured TI conditions given in table
\ref{tab:CapeWindMeasurements}, the surface roughness and prescribed
surface temperature change were adjusted through an initial
trial-and-error process.  The final surface roughness of $z_0$=0.0005m
used for the stable offshore conditions (see table \ref{tab:z0tempparam})
matched the roughness used in the neutral 15m/s and unstable 5m/s and
10m/s cases from the previous study \cite{cheung2020large}.  This
value of surface roughness is consistent with the measurements from
the North Sea \cite{taylor2001dependence}, which found values of $z_0$
ranging from $5 \times 10^{-5}$m to $5\times 10^{-3}$m.

The values of the prescribed temperature change at the ocean surface
is also given in table \ref{tab:z0tempparam}.  As the wind
speed increases, a larger temperature decrease was used to maintain
similar levels of stable stratification, as expected from
Monin-Obukhov similarity theory.
